---
title: "Project KillR"
output: 
  flexdashboard::flex_dashboard:
    orientation: rows
    theme: yeti
    navbar:
      - { title: "About", href: "https://dataplanes.org" }
    source_code: https://github.com/lhehnke/project-killR

---

<style>
  
  @import url(https://fonts.googleapis.com/css?family=Open+Sans:300);
.navbar-inverse {
  background-color: #7d3945;
  border-color: #7d3945;
}

</style>

```{r, include = FALSE}
# Load flexdashboard
library(flexdashboard)

# Install and load packages using pacman
if (!require("pacman")) install.packages("pacman")
library(pacman)

p_load(DT, highcharter, leaflet, lubridate, kableExtra, plotly, tidyverse)

# Set directory
MAIN_DIR <- rprojroot::find_rstudio_root_file()

# Import data
killers_df <- readRDS(file = paste(MAIN_DIR, "serial_killers_data.rds", sep = "/")) 
```

Data {data-icon="fa-chart-line"}
=======================================================================

Row
-----------------------------------------------------------------------

### This table contains data on `r nrow(killers_df)` serial killers from `r killers_df %>% filter(!is.na(country)) %>% distinct(country) %>% nrow()` countries that was scraped from [Wikipedia](https://de.wikipedia.org/wiki/Liste_von_Serienmördern) in February 2018 (click [here](https://github.com/lhehnke/serial-killers) for the replication script). You can search the table, sort the columns, and filter the entries.

```{r}
killers_df %>% 
  select(name, pseudonym, sex, country, victims_proven, victims_suspected) %>%
  arrange(name, -victims_proven) %>%
  rename(Killer = name, Alias = pseudonym, Sex = sex, Country = country, "Victims (proven)" = victims_proven, "Victims (suspected)" = victims_suspected) %>% 
  DT::datatable(filter = "bottom", extensions = "Buttons", options = list(dom = "Bfrtip", pageLength = 100, initComplete = JS(
    "function(settings, json) {",
    "$(this.api().table().header()).css({'background-color': '#a9d3e1', 'color': '#000'});",
    "}"), 
    buttons = list("copy", list(extend = "collection", buttons = list(
    list(extend = "csv", filename = "Serial_killeRs", title = "Project KillR: International serial killers from 1435-2013"), 
    list(extend = "excel", filename = "Serial_killeRs", title = "Project KillR: International serial killers from 1435-2013"), 
    list(extend = "pdf", filename = "Serial_killeRs", title = "Project KillR: \nInternational serial killers from 1435-2013")), text = "Download"))))
```

Killers {data-icon="fa-user-secret"}
=======================================================================

Row
-----------------------------------------------------------------------

### Killers {.value-box}

```{r}
killers <- nrow(killers_df)
valueBox(killers, icon = "fa-user-secret") 
```

### Countries {.value-box}

```{r}
countries <- killers_df %>% 
  filter(!is.na(country)) %>%
  distinct(country) %>%
  nrow()
valueBox(countries, icon = "fa-globe")
```

### Time span {.value-box}

```{r}
to_from <- killers_df %>%
  mutate(year_min = min(active_from, na.rm = TRUE), year_max = max(active_to, na.rm = TRUE)) %>%
  unite_("year_range", c("year_min","year_max"), sep = "-") %>%
  slice(1) %>%
  pull()
valueBox(to_from, icon = "fa-history")
```

Row {.tabset .tabset-fade}
-----------------------------------------------------------------------

### Who are the deadliest serial killers?

```{r}
# Plot killers with most proven victims
p <- killers_df %>%
  group_by(name) %>%
  summarise(victims_proven = sum(victims_proven, na.rm = TRUE)) %>% 
  top_n(10, victims_proven) %>%
  ungroup() %>%
  mutate(name = reorder(name, victims_proven)) %>%
  rename(Killer = name, `Victims (proven)` = victims_proven) %>%
  ggplot(mapping = aes(x = Killer, y = `Victims (proven)`)) +
  geom_bar(stat = 'identity', width = 0.5, alpha = 0.8, fill = "#7d3945", color = "#7d3945") +
  labs(x = "", y = "") + ylim(0, 250) + coord_flip() + theme_minimal()

ggplotly(p)
```

### Are serial killers mainly male or female?

```{r}
# Number of killers by sex
p <- killers_df %>%
  group_by(sex) %>%
  dplyr::count(sort = TRUE) %>%
  rename(Sex = sex, Count = n) %>%
  ggplot(mapping = aes(x = Sex, y = Count)) +
  geom_bar(aes(fill = Sex), stat = 'identity', width = 0.5, alpha = 0.8, color = "#7d3945") +
  scale_fill_manual("Sex", values = c("female" = "#a04958", "male" = "#5a2932")) +
  labs(x = "", y = "") + theme_minimal()

ggplotly(p)
```

### Who are the deadliest male and female killers?

```{r}
# Plot killers with most proven victims by sex
p <- killers_df %>%
  group_by(sex) %>%
  top_n(10, victims_proven) %>%
  ungroup() %>%
  mutate(name = reorder(name, victims_proven)) %>%
  rename(Killer = name, `Victims (proven)` = victims_proven, Sex = sex) %>%
  ggplot(mapping = aes(x = Killer, y = `Victims (proven)`)) +
  geom_bar(aes(fill = Sex), stat = 'identity', width = 0.5, alpha = 0.8, color = "#7d3945") +
  scale_fill_manual(name = "", values = c("female"= "#a04958", "male" = "#5a2932")) + 
  labs(x = "", y = "") + ylim(0, 250) + coord_flip() + theme_minimal()

ggplotly(p)
```

### Where do they come from?

```{r}
# Plot deadliest countries with information on sex of killers
p <- killers_df %>%
  filter(!is.na(country)) %>%
  group_by(sex) %>%
  dplyr::count(country, sort = TRUE) %>%
  ungroup() %>%
  top_n(10, n) %>%
  ungroup() %>%
  mutate(country = reorder(country, n)) %>%
  rename(Country = country, Count = n, Sex = sex) %>%
  ggplot(mapping = aes(x = Country, y = Count)) +
  geom_bar(aes(fill = Sex), stat = 'identity', width = 0.5, alpha = 0.8, color = "#7d3945") +
  scale_fill_manual(name = "", values = c("female"= "#a04958", "male" = "#5a2932")) + 
  labs(x = "", y = "") + ylim(0, 250) + coord_flip() + theme_minimal()

ggplotly(p)
```

Victims {data-icon="fa-skull"}
=======================================================================

Row
-----------------------------------------------------------------------

### Proven victims {.value-box}

```{r}
victims_prov <- killers_df %>% 
  summarize(sum(victims_proven, na.rm = TRUE)) %>%
  pull()
valueBox(victims_prov, icon = "fa-skull")
```

### Suspected victims {.value-box}

```{r}
victims_sus <- killers_df %>%
  mutate(victims_suspected = as.numeric(victims_suspected)) %>%
  summarize(sum(victims_suspected, na.rm = TRUE)) %>%
  pull()
valueBox(victims_sus, icon = "fa-question")
```

### Time span {.value-box}

```{r}
to_from <- killers_df %>%
  mutate(year_min = min(active_from, na.rm = TRUE), year_max = max(active_to, na.rm = TRUE)) %>%
  unite_("year_range", c("year_min","year_max"), sep = "-") %>%
  slice(1) %>%
  pull()
valueBox(to_from, icon = "fa-history")
```

Row
-----------------------------------------------------------------------

### Who killed more victims: Male or female serial killers?

```{r}
# Plot killings by sex of killers
## TODO: Sex shows twice in plotly graph. Remove once.
p <- killers_df %>%
  group_by(sex) %>%
  summarise(victims_proven = sum(victims_proven, na.rm = TRUE)) %>% 
  top_n(10, victims_proven) %>%
  ungroup() %>%
  mutate(sex = reorder(sex, victims_proven)) %>%
  rename(Sex = sex, `Victims (proven)` = victims_proven) %>%
  ggplot(mapping = aes(x = Sex, y = `Victims (proven)`)) +
  geom_bar(aes(fill = Sex), stat = 'identity', width = 0.5, alpha = 0.8, color = "#7d3945") +
  scale_fill_manual(name = "", values = c("female"= "#a04958", "male" = "#5a2932")) + 
  labs(x = "", y = "Proven victims") + theme_minimal()

ggplotly(p)
```

### Who killed the most victims? {data-width=440} 

```{r}
killers_df %>% 
  select(name, sex, victims_proven, victims_suspected) %>%
  arrange(-victims_proven, name) %>%
  rename(Killer = name, Sex = sex, "Victims (proven)" = victims_proven, "Victims (suspected)" = victims_suspected) %>%
  kable(escape = FALSE) %>%
  kable_styling(bootstrap_options = c("striped", "hover"))
```

Row
-----------------------------------------------------------------------

### In which countries were the most victims killed? 

```{r}
# Plot killing by country
p <- killers_df %>%
  filter(!is.na(country)) %>%
  group_by(country, sex) %>%
  summarise(victims_proven = sum(victims_proven, na.rm = TRUE)) %>% 
  ungroup() %>%
  top_n(10, victims_proven) %>%
  mutate(country = reorder(country, victims_proven)) %>%
  rename(Country = country, `Victims (proven)` = victims_proven, Sex = sex) %>%
  ggplot(mapping = aes(x = Country, y = `Victims (proven)`)) +
  geom_bar(aes(fill = Sex), stat = 'identity', width = 0.5, alpha = 0.8, color = "#7d3945") +
  scale_fill_manual(name = "", values = c("female"= "#a04958", "male" = "#5a2932")) + 
  labs(x = "", y = "") + ylim(0, 2000) + coord_flip() + theme_minimal()

ggplotly(p)
```

Timeline {data-icon="fa-history"}
=======================================================================

Row
-----------------------------------------------------------------------

```{r, include = FALSE}
# Calculate timeline
timeline <- killers_df %>%
  select(name, active_from, active_to) %>%
  na.omit() %>%   
  gather(status, years, -name) %>%
  select(-status) %>%
  group_by(name) %>% 
  complete(name, years = full_seq(years, 1)) %>%
  ungroup() %>%
  count(years) %>%
  rename(Year = years, Count = n) %>%
  ungroup() %>% 
  mutate(Year = lubridate::ymd(Year, truncated = 2L)) %>% 
  complete(Year = seq.Date(min(Year), max(Year), by = "year")) %>%
  mutate(Year = year(Year)) %>%
  mutate(Count = replace(Count, is.na(Count), 0))
```

### The first serial killer appeared in `r timeline %>% summarize(min(Year)) %>% pull()`. In 2013, `r timeline %>% filter(Year == 2013) %>% pull(Count)` killers were still active, while the peak was in 1990 with `r timeline %>% summarize(max(Count)) %>% pull()` killers. Hover over the timeline to display the count for each year. {data-width=740}

```{r}
# Plot number of killers active
timeline %>%
  hchart("line", hcaes(x = "Year", y = "Count"), name = "Killers active", color = "#7d3945")
```

### This table shows the number of active serial killers for each year from `r timeline %>% summarize(min(Year)) %>% pull()` to `r timeline %>% summarize(max(Year)) %>% pull()`. {data-width=300}

```{r}
timeline %>%
  arrange(-Count, Year) %>%
  kable(escape = FALSE) %>%
  kable_styling(bootstrap_options = c("striped", "hover"))
```  

Locations {data-icon="fa-map-marker-alt"}
=======================================================================

Row
-----------------------------------------------------------------------

```{r, include = FALSE}
# Select columns to merge
victims_merge <- killers_df %>%
  filter(!is.na(country)) %>%
  group_by(country) %>%
  mutate(victims_suspected = as.numeric(victims_suspected)) %>%
  summarize_at(vars(victims_proven = victims_proven, victims_suspected = victims_suspected), sum, na.rm = TRUE) %>%
  select(country, victims_proven, victims_suspected) %>% 
  mutate_at(vars(victims_proven, victims_suspected), list(~ ifelse(is.na(.), 0, .))) %>%
  mutate(country = gsub("United States", "United States of America", country)) %>%
  mutate(victims_share = victims_proven/sum(victims_proven)) # calculate share

# Download map and merge data
world_map <- get_data_from_map(download_map_data("custom/world-eckert3"))

map_killings <- world_map %>%
  left_join(victims_merge, by = c("name" = "country")) %>% 
  mutate_at(.vars = vars(victims_proven, victims_suspected), .funs = funs(ifelse(is.na(.), 0, .)))
```

### This interactive map shows the number of proven victims for each country. {data-width=740}

```{r}
# Plot killings by country
hcmap("custom/world-eckert3", data = map_killings, value = "victims_proven", 
      joinBy = "hc-a2", name = "Proven victims",
      dataLabels = list(enabled = TRUE, format = "{point.name}"),
      borderColor = "#FAFAFA", borderWidth = 0.1) %>%
  hc_colorAxis(minColor = "#D4BDC1", maxColor = "#7d3945") %>%
  hc_mapNavigation(enabled = TRUE)
```

### This table provides an overview of the countries with the highest number of serial killers from `r timeline %>% summarize(min(Year)) %>% pull()` to `r timeline %>% summarize(max(Year)) %>% pull()`. {data-width=300} 

```{r}
killers_df %>%
  select(name, pseudonym, country) %>%
  dplyr::count(country, sort = TRUE) %>%
  ungroup() %>%
  mutate(country = reorder(country, n)) %>%
  filter(!is.na(country)) %>%
  rename(Country = country, Count = n) %>%
  kable(escape = FALSE) %>%
  kable_styling(bootstrap_options = c("striped", "hover"))
```  
